Ab Initio Effective One-Electron Potential Operators. II. Applications for Exchange-Repulsion Energy in Effective Fragment Potentials

### 2019-12-21 papers/3_oep_technique.Pauli/paper-r3.tex
Figures in Paper 3 updated

\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

@@

\begin{abstract}
In Paper~I, the effective one\hyp{}electron
potentials (OEP) method was introduced and demonstrated as an
efficient approach
to reduce the computational cost
of evaluation of the charge-transfer interaction energy
within the effective fragment potential method (EFP2)
by an average factor of 20, making it no longer a bottleneck
in EFP2\hyp{}based simulations of complex systems.
Here, the OEP technique 
is used to enhance computational efficiency in evaluating the
exchange\hyp{}repulsion EFP2 interaction energy
by redefining the first\hyp{}order repulsive term of 
Murrell~et al. [Murrell et al., Proc. R. Soc. Lond. A 284, 566 (1965)] through
the extended density fitting in incomplete auxiliary basis.
In the proposed approach the evaluation kinetic energy integrals is no
longer required and the computational cost can be reduced roughly by a factor
of 1.5 as compared to the original EFP2 formulation.
\end{abstract}

@@

The exchange\hyp{}repulsion (EXR) interaction energy in the
second generation of the
effective fragment potentials (EFP2) method\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
describes the non\hyp{}electrostatic repulsion
between two isolated and unperturbed Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
wavefunctions in an interacting
bi\hyp{}molecular complex.
The EXR EFP2 term was derived from the intermolecular perturbation theory
of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
up to the second order with respect to the wavefunction overlap,
and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:exr-efp2}
 E^{\rm Ex-Rep} \approx
 E^{\rm Ex} +
 E^{\rm Rep}(S^{-1}) + 
 E^{\rm Rep}(S^{-2}) \;.
\end{equation}
%
In the above equation,
$E^{\rm Ex}$ is the exchange energy,\cite{Jensen.JCP.1996}
%
\begin{equation} \label{e:exc-efp2}
 E^{\rm Ex} \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} 
 \sqrt{\frac{-2\ln{\vert S_{ij}\vert}}{\pi}} \frac{S^2_{ij}}{r_{ij}} \;,
\end{equation}
%
whereas the first\hyp{} and second\hyp{}order repulsion terms
are accordingly~\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{multline} \label{e:rep-efp2.s1}
 E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} \\
 \times
 \left\{
 \sum_{k\in A}^{\rm Locc} F_{ik}^A S_{kj} 
+\sum_{l\in B}^{\rm Locc} F_{jl}^A S_{li} 
 -2T_{ij}
 \right\}
\end{multline}
%
and 
%
\begin{multline} \label{e:rep-efp2.s2}
 E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
 \\ \times 
 \Bigg\{
 \sum_{x\in A}^{\rm At}  \frac{-Z_x}{r_{xb}}
+\sum_{y\in B}^{\rm At}  \frac{-Z_y}{r_{ya}} 
+\sum_{k\in A}^{\rm Locc} \frac{   2}{r_{jk}}
+\sum_{l\in B}^{\rm Locc} \frac{   2}{r_{il}}
-r_{ij}^{-1}
 \Bigg\} \;.
\end{multline}
%
The indices $i$, $j$, $k$ and $l$ label the localized occupied orbitals (LMO's, denoted by `Locc')
located at their charge centroids ${\bf r}_{i(j)}$,
whereas $x$ and $y$ label atomic nuclei (denoted by `At') with atomic numbers 
$Z_{x(y)}$ located at ${\bf r}_{x(y)}$. The relative distances are
defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
It is known that evaluating Eq.~\eqref{e:exr-efp2}
is remarkably efficient and gives usually very accurate
estimates of the reference EXR energies according to the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
The computational cost to evaluate
the exchange and second\hyp{}order terms
is much smaller than that of the first\hyp{}order term,
which requires evaluation of not only the one\hyp{}electron overlap,
but also the kinetic
energy integrals in a space of occupied molecular orbital (MO) basis.

@@

           W^A_{ij} 
         + W^B_{ji} \\

@@

 \BraKet{ij}{kk} &\equiv \sum_{\mu\in A} 

@@

 \BraKet{ik}{jk} &\equiv \sum_{\mu\in A} 

@@

  \hat{v}_{ik} \equiv \int d{\bf r} \Ket{{\bf r}} 

@@

 V^A_{ij} + 

@@

In this procedure, the required OEP matrix elements can be calculated from

@@

     = \sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} \\

>>>>

\input{header.tex}
\begin{document}

@@

\begin{abstract}
In Paper~I, the effective one\hyp{}electron
potentials (OEP) method was introduced and demonstrated as an
efficient approach
to reduce the computational cost
of evaluation of the charge-transfer interaction energy
within the effective fragment potential method (EFP2)
by an average factor of 20, making it no longer a bottleneck
in EFP2\hyp{}based simulations of complex systems.
Here, the OEP technique 
is used to enhance computational efficiency in evaluating the
exchange\hyp{}repulsion EFP2 interaction energy
by redefining the first\hyp{}order repulsive term of 
Murrell~et al. [Murrell et al., Proc. R. Soc. Lond. A 284, 566 (1965)] through
the extended density fitting in incomplete auxiliary basis.
In the proposed approach the evaluation of the kinetic energy integrals is no
longer required and the computational cost can be reduced roughly by a factor
of 1.5 as compared to the original EFP2 formulation.
\end{abstract}

@@

The exchange\hyp{}repulsion (EXR) interaction energy in the
second generation of the
effective fragment potentials (EFP2) method\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
describes the non\hyp{}electrostatic repulsion
between two isolated and unperturbed Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
wavefunctions in an interacting
bi\hyp{}molecular complex.
The EXR EFP2 term was derived from the intermolecular perturbation theory
of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
up to the second order with respect to the wavefunction overlap,
and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:exr-efp2}
 E^{\rm Ex-Rep} \approx
 E^{\rm Ex} +
 E^{\rm Rep}(S^{-1}) + 
 E^{\rm Rep}(S^{-2}) \;.
\end{equation}
%
In the above equation,
$E^{\rm Ex}$ is the exchange energy,\cite{Jensen.JCP.1996}
%
\begin{equation} \label{e:exc-efp2}
 E^{\rm Ex} \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} 
 \sqrt{\frac{-2\ln{\vert S_{ij}\vert}}{\pi}} \frac{S^2_{ij}}{r_{ij}} \;,
\end{equation}
%
whereas the first\hyp{} and second\hyp{}order repulsion terms
are accordingly~\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{multline} \label{e:rep-efp2.s1}
 E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} \\
 \times
 \left\{
 \sum_{k\in A}^{\rm Locc} F_{ik}^A S_{kj} 
+\sum_{l\in B}^{\rm Locc} F_{jl}^A S_{li} 
 -2T_{ij}
 \right\}
\end{multline}
%
and 
%
\begin{multline} \label{e:rep-efp2.s2}
 E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
 \\ \times 
 \Bigg\{
 \sum_{x\in A}^{\rm At}  \frac{-Z_x}{r_{xb}}
+\sum_{y\in B}^{\rm At}  \frac{-Z_y}{r_{ya}} 
+\sum_{k\in A}^{\rm Locc} \frac{   2}{r_{jk}}
+\sum_{l\in B}^{\rm Locc} \frac{   2}{r_{il}}
-r_{ij}^{-1}
 \Bigg\} \;.
\end{multline}
%
The indices $i$, $j$, $k$ and $l$ label the localized occupied orbitals (LMO's, denoted by `Locc')
located at their charge centroids ${\bf r}_{i(j)}$,
whereas $x$ and $y$ label atomic nuclei (denoted by `At') with atomic numbers 
$Z_{x(y)}$ located at ${\bf r}_{x(y)}$. The relative distances are
defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
$S_{ij}$ and $T_{ij}$ are the overlap and kinetic\hyp{}energy
integrals, respectively, whereas $F_{ik}^X$ are the Fock matrix elements
of unperturbed monomer $X$, all defined in LMO basis.
It is known that evaluating Eq.~\eqref{e:exr-efp2}
is remarkably efficient and gives usually very accurate
estimates of the reference EXR energies according to the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
The computational cost to evaluate
the exchange and second\hyp{}order terms
is much smaller than that of the first\hyp{}order term,
which requires evaluation of not only the one\hyp{}electron overlap,
but also the kinetic
energy integrals in a space of occupied molecular orbital (MO) basis.

@@

         - W^A_{ij} 
         - W^B_{ji} \\

@@

 \BraKet{ij}{kk} &\equiv - \sum_{\mu\in A} 

@@

 \BraKet{ik}{jk} &\equiv - \sum_{\mu\in A} 

@@

  \hat{v}_{ik} \equiv - \int d{\bf r} \Ket{{\bf r}} 

@@

 -W^A_{ij} + 

@@

In this procedure, $R_{\alpha\gamma}= \iint d{\bf r}_1 d{\bf r}_2 \frac{\phi_{\alpha}^*({\bf r}_1) \phi_{\beta}({\bf r}_2) }{\lvert {\bf r}_1 - {\bf r}_2 \rvert}$, ${\bf S}$ is the matrix of overlap integrals in AO basis
and the required OEP matrix elements can be calculated from

@@

     = -\sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} \\
